
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% 
\section{Matching the ND Requirements for DUNE/LBNF} 
\label{cdrsec:detectors-nd-ref-fgt-req}

The scope of the this section addresses how the the reference NND
described in Section~\ref{cdrsec:detectors-nd-ref} meets the
requirements of the oscillation studies described in
Section 3.8.2 of~\cite{cdr-vol-2} %~\ref{sec:physics-lbnosc-beamnd-req} 
and the short-baseline precision measurements described in Sections 6.1.2
and 6.1.3 of the same document. %~\ref{sec-nd-sbp} and ~\ref{sec-nd-np}.  
First,
we present the oscillation related systematics; the systematics
affecting the precision measurements and new-physics search program
follows.


%\subsection{Matching ND Requirements for the Oscillation Analyses} 
\subsection{Oscillation Analyses} 
\label{cdrsec:detectors-nd-ref-fgt-req-oscl}

The Table 3.8   % Anne updated 6/20
%~\ref{tab:nuesysts} 
in Section 3.6.2 of~\cite{cdr-vol-2}
%~\ref{sec:physics-lbnosc-beamnd-req} 
presents a conservative projection of systematic errors affecting
$\nu_e$ appearance. The FGT measurement alleviates the systematic
errors as enumerated below:
\begin{itemize}
    \item {\bf Beam $\nu_e$:} The FGT will offer an event-by-event
      measurement of the beam $\nu_e$ via the identification of the
      emergent $e^-$ in STT while rejecting the $\pi^0 \rightarrow
      \gamma \rightarrow e$ background via the determination of the
      missing-$P_T$ vector to a high degree.  In a 5-year neutrino-run
      (focus-positive) the FGT will accumulate $500,000$ $\nu_e$
      with $\simeq 55\%$ efficiency and $\geq 95\%$ purity.
      The resulting $\nu_e / \nu_\mu$ ratio will be determined to
      $\leq 1\%$ precision. Furthermore, by constraining the sources
      of $\nu_e$ ($\mu^+$, $K^+$ and $K^0_L$), the FGT will predict
      the the ratio of $\nu_e$ and $\nu_\mu$ spectra at the far
      detector (FD) with respect to the near detector (ND) as a
      function of the neutrino energy, FD/ND ($E_\nu$).

    \item {\bf Beam $\bar\nu_e$:} Although the FD does not distinguish
      $\nu_e$ from $\bar\nu_e$, the FGT will accurately measure the
      beam $\bar\nu_e$ by identifying the emergent $e^+$ in the STT
      with efficiency and purity similar to those for $\nu_e$.  (We
      point out that the dominant kaonic source of $\bar\nu_e$ is
      $K^0_L$; the neutrino spectra from $K^0_L$ are different from
      those of $K^+$ at the FD.)  In a 5-year neutrino-run, the FGT
      will accumulate 40,000 $\bar\nu_e$ and $5\times 10^6$
      $\bar\nu_\mu$ samples, providing a precise FD/ND prediction for
      the antineutrino to $\simeq 1\%$.  These constraints will be
      even more valuable during the antineutrino run (focus-negative)
      where the wrong-sign backgrounds are larger.

    \item {\bf Cross Sections:} First, the FGT will 
      measure the absolute flux, via $\nu$-e scattering,
      to $\simeq 2.5\%$ precision. Second, using radiator targets,
      the FGT will measure exclusive channels, such as
      quasi-elastic, resonance, coherent-mesons and the inclusive DIS
      channel with unparalleled precision. Since Argon is the
      far detector nuclear target, a set of various nuclear targets will
      allow translation of the cross-section measurements to $\nu$-Ar
      scattering.

    \item {\bf Nuclear Effects:} The FGT will employ a suite of
      nuclear targets including Ar-gas in pressurized tubes, a thin
      solid calcium target (which has the same A=40 at Ar), a
      C-target, etc.  Specifically, the number of $\nu$-Ar interaction
      will be 10 times larger than that expected in a \ktadj{40} FD,
      without oscillations. Additionally, comparisons of calculations
      of elastic and inelastic interactions in Ar versus Ca, including
      FSI effects, indicate negligible differences between the two
      targets. Thus, the combination of the Ar and Ca targets will
      provide a strong constraint on the nuclear effects from both
      initial and final state interactions.  Finally, FGT's ability to
      isolate $\nu$ ($\bar\nu$) off free-hydrogen, via subtraction of
      hydrocarbon and carbon targets, will provide a 
      {\em model-independent} measurement of nuclear effects.

    \item {\bf Hadronization:} A notable strength of FGT is to
      identify the yield of $\pi^0$ $separately$ in neutral-current
      (NC) and charged-current (CC) interactions. (The estimated
      $\pi^0$ detection efficiency is $\simeq 50\%$.) In addition, FGT
      will determine the yields of $\pi^-$ and $\pi^+$, the dominant
      backgrounds to the $\nu_\mu$ and $\bar\nu_\mu$
      disappearance. Finally, the measurement of the composition,
      energy, and angle of the hadronic jet will provide a tight
      constraint on the overall hadronization models.

    \item {\bf Energy Scale:} The average density of about
      0.1\,g/cm$^3$ (close to that of liquid hydrogen) makes the STT
      mostly transparent to secondary particles --- the entire STT
      length is equivalent to $\sim 1.4 X_0$ including the various
      nuclear targets. As a result we will be able to measure
      accurately the 4-momenta of secondary particles as well as the
      missing-$P_T$ vector in the CC processes.  This redundant
      missing-$P_T$ vector measurement provides a most important
      constraint on the neutrino and antineutrino energy scales. The
      capabilities of this kinematic analysis have been demonstrated
      by the NOMAD experiment (with similar density and B field).  In
      addition, measurements of exclusive topologies like
      quasi-elastic, resonance and coherent meson production offer
      additional constraints on the nuclear effects affecting the
      neutrino energy scale (see Section 6.1.2 of~\cite{cdr-vol-2}).  %~\ref{sec-nd-sbp}).  
The requirements on the muon energy scale
      ($< 0.2\%$) and total hadron energy scale ($< 0.5\%$)
      uncertainties have been already achieved by the NOMAD
      experiment. The muon energy scale is calibrated with the mass
      constraint from the $K^0$ reconstructed in STT and was
      statistics limited in NOMAD. The total hadron energy scale is
      calibrated using the $y_{Bj}$ distribution in different energy
      bins. Compared to NOMAD, FGT will have $\times 100$ higher
      statistics and $\times 10$ higher granularity.
\end{itemize}

In summary the FGT will accurately quantify all four neutrino species and
predict the ratio FD/ND for them.  It will measure the 4-momenta of
the outgoing hadrons composing the hadronic jets in a variety of
nuclear targets, in essence proving a data-driven \textit{event generator}
which can be applied to the FD.

Since the FGT is based upon a different technology than the FD, it cannot
account for effects of LAr reconstruction inefficiencies in the
FD. The corresponding cancellation could be achieved only with an
identical ND, which to some extent is an ill-defined concept due to a
number of factors including size, beam profile and composition, rates, 
etc. However, given the detailed program to calibrate
LAr detectors in test beams and  multiple neutrino
experiments employing LAr detectors, by the time DUNE becomes
operational the reconstruction of particles in LAr will likely be well
understood. Finally, Section~\ref{sec:detectors-nd-alt} outlines
the enhancement of the ND complex via the placement of complementary
LAr detector(s) upstream of the FGT.

%\subsection{Matching ND Requirements for the Short Baseline Precision Measurements and Searches} 
\subsection{Short Baseline Precision Measurements and Searches} 
\label{cdrsec:detectors-nd-ref-fgt-req-sbp}


Sections 6.1.2
%~\ref{sec-nd-sbp} 
and 6.1.3 of~\cite{cdr-vol-2}
 %~\ref{sec-nd-np} 
summarize a rich physics program at the near site providing a
generational advance in precision measurements and sensitive searches.
This short-baseline physics program and the long-baseline oscillation
analyses share similar detector requirements and offer a deep synergy
and mutual feedback.  The reference FGT meets the requirements of the
short-baseline studies as briefly outlined below:
\begin{itemize}
    \item {\bf Resolution:} The FGT is designed to have an order of
      magnitude higher granularity than NOMAD, the most precise, high
      statistics neutrino experiment. The corresponding improvements
      include better tracking, electron/positron ID through TR, $dE/dx$
      measurement providing hadron-ID, 4$\pi$ calorimetry, 4$\pi$ muon
      coverage and a larger transverse area for event containment.
    \item {\bf Statistics:} The \MWadj{1.2} neutrino source at LBNF will
      offer a factor of 100 enhancement in statistics compared to
      NOMAD. The program of measuring $\nu$ and $\bar\nu$ interactions
      in a set of nuclear targets, including Ar and H, will enhance
      the physics potential of precision measurements and searches.
\end{itemize}
Sensitivity studies to the salient precision measurements and searches
can be found in~\cite{Adams:2013qkq, DPR}.


\subsection{Future Tasks to Quantify the Systematic Errors}
\label{cdrsec:detectors-nd-ref-fgt-req-future} 

%We need to undertake three outstanding tasks to quantify the
%systematic errors in oscillation studies and precision measurement
%program:
To quantify the systematic errors in oscillation studies and precision measurement
program, three tasks are still outstanding:\begin{itemize}
    \item {\bf Geant4 Simulation:} A Geant4 simulation of the FGT is
      needed to confirm and correct the projected systematic errors
      and the salient sensitivity studies.
     \item {\bf Event Reconstruction:} A program to reconstruct tracks
       in STT and to match the information from different
       subdetectors is needed to identify secondary particles.
     \item {\bf Translating ND-Measurements to FD:} A concerted effort
       needs to be mounted to transfer the precision measurements in
       ND to the far detector.
\end{itemize}    
The DUNE collaboration plans to pursue these issues with high priority
in the coming years before CD-2.
