%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Near Detector Systems}
\label{sec:detectors-nd-ref-ov}

\fixme{Anne changed the first part of the intro a bit 4/28 to be more standard}
This chapter describes the reference design of the DUNE Near Detector Systems (NDS). The scope includes the design, procurement, fabrication, testing, delivery and installation of the NDS components: 

\begin{itemize}
\item Fine-Grained Tracker (FGT) near neutrino detector (NND);
\item Beamline Measurement System (BLM);  
\item NDS Data Acquisition system (DAQ).  
\end{itemize}
Detailed descriptions of the systems and components in the NDS are provided in \anxndref. 


The concept and design of the reference DUNE-ND evolved, first,  from the experience gained 
from MINOS, the first generation of long-baseline neutrino experiment at Fermilab, 
and NOvA, the second generation experiment, and, next, from the high resolution NOMAD detector at CERN 
and the T2K detector at KEK. MINOS and NOvA experiments employ functionally `identical' detectors which 
fulfill the mission of these experiments, given the statistics and resolution of the respective far detectors. 
DUNE, the third generation experiment at Fermilab, has a much more ambitious set of goals than 
either MINOS or NOvA,  and the effects that we seek to illuminate are
considerably more subtle: 
discovery of CP-violation, discovery of mass hierarchy, and a search for 
physics beyond PMNS with unparalleled precision. Concomitantly DUNE will have a more intense 
neutrino source and a much higher resolution, and a more massive FD.  
To meet the systematic precision needed to fulfill these goals,
the ND must thoroughly characterize the beam, composed of muon-neutrino and -antineutrino 
and electron-neutrino and -antineutrino at the source. Additionally, it must also precisely measure 
the cross-sections and the particle-yields of various processes that compose  neutrino events. 
By particle-yields we mean the multiplicity and momentum distributions of pions, kaons, protons, etc., 
produced in the hadronic jet, that constitute the backgrounds to the oscillation signal -- appearance or 
disappearance. An ``identical ND", like those in MINOS or NOvA with the experience gained therein, 
cannot accomplish these requirements. 

As pointed out in Chapter~\ref{ch:physics-nd} of Volume-2, because 
not a single spectrum at FD is identical to that at the ND the concept of `identical' detectors 
is an over-simplification.  Furthermore the onus of precisely quantifying the neutrino 
source and the cross-sections, including the hadronic composition, impels one toward 
a high resolution ND. The NOMAD experience evinces that a high resolution detector,  
capable of measuring $e^{-}$, $e^{+}$, $\mu^{-}$, $\mu^{+}$, $\pi^{-,+,0}$, proton and $K^{0}$, 
would  partially meet the challenges of DUNE -- the detector must be augmented to measure, and thereby 
precisely model,  the neutrino-nuclear effects. The reference DUNE-ND, is, thus,  a next generation 
of near-detector concept as compared to the T2K experiment. Such a detector would also 
vastly enrich the physics potential of the DUNE/LBNF program. 
Finally, a complementary LAr-detector(s) upstream of the high resolution ND would enhance 
the capability of the ND complex. 


The reference DUNE-ND, a  Fine-Grained Tracker (FGT) near detector,  consists of a straw-tube
tracking detector (STT) and electromagnetic calorimeter (ECAL) inside of a 0.4-T
dipole magnet. In addition, Muon Identifiers (MuIDs) are located in the
steel of the magnet, as well as upstream and downstream of the STT. The FGT
is designed to make precision measurements of the neutrino fluxes, 
cross sections, signal rates and background rates at the percent level. 

The Beamline Measurement System (BLM) will be located in the region of the Absorber Complex at 
the downstream end of the decay region to measure the muon fluxes from hadron decay. The 
absorber itself is part of the LBNF Beamline. 
The BLM is intended to determine the neutrino fluxes and spectra
and to monitor the beam profile on a spill-by-spill basis, and will operate for the life of the
experiment. 

The Near Detector System Data Acquisition system (NDS-DAQ) collects raw data from each NDS detector's
individual DAQ, issues triggers, adds precision timing 
data from a global positioning system (GPS), and builds events. 
The NDS-DAQ is made up of three parts: NDS Master DAQ (NDS-MDAQ), the Beamline Measurements 
DAQ (BLM-DAQ) and the Near Neutrino Detector DAQ (NND-DAQ).



%\begin{cdrtable}[short]{cc}{label}{long}  %The third argument (reads {cc}) can use c, l, r or p{some length} 
%%%% but please do not include lines like “|c|l|l|”. It CAN look like {cll} or {llp{3cm}}, for instance.
%Header Column1 & Header Column 2 \\ \toprowrule
%Row 1 & First \\ \colhline
%Row 2 & Second \\ \colhline
%Row 3 & Third \\
%\end{cdrtable}


