\chapter{Project Management}
\label{ch:detectors-pm}

The international DUNE experiment is managed through two
tightly-coupled structures: the DUNE Science Collaboration and the
DUNE Project Office.  The DUNE Science Collabora tion is responsible
for the scientific and experimental strategy, while the DUNE Project
Office is responsible for the implementation of this strategy within
the context of the available international resources.

\section{DUNE Collaboration}

The DUNE Science Collaboration is the organized structure that brings
together the members of the international science community interested
in participating in the DUNE experiment.  The Collaboration defines
the scientific goals of the experiment and subsequently the
requirements on the experimental facilities needed to achieve these
goals.  The Collaboration also provides the scientific effort required
for the design of the detectors and facilities, operation of the
experiment, and analysis of the collected d ata.

An Institutional Board (IB), composed of one representative from each
of the member institutions, has responsibility for Collaboration
governance.  The IB has final authority over Collaboration membership
issues and defines requirements for inclusion of individuals within
the DUNE authorship list. The IB is also responsible for establishing
and monitoring the process through which the co-spokespeople are
selected to serve as leaders of the collaboration.

The DUNE Executive Committee (EC) advises the co-spokespeople in
dealing with strategic issues affecting the experiment such as
technological design choices and the distribution of project
responsibilities among the participating institutions.  Membership of
the EC includes the co-spokespeople, DUNE Project Office leaders, IB
chair, and five additional Collaboration members (three elected IB
representatives and two additional members selected by the
co-spokespeople).

\section{DUNE Project Office}

The DUNE Project is organized as an international entity that manages
all contributions to the design, construction, installation, and
commissioning of the DUNE near and far detectors.  The U.S. DOE-funded
portion of the Project is contained within the international entity,
and the management of this portion is a direct responsibility of the
international DUNE project team.  The entity will be run as an
international project matching DOE requirements. This will imply
maintaining a full cost and schedule for the entire project, from
which the DOE-funded portion can be extracted and monitored in a
manner that satisfies DOE reporting requirements.

The Project will have direct control over DOE project funds and any
common fund (e.g., construction and operation) collected from the
U.S. and international stakeholders. Unless requested otherwise,
Project funding obtained through other agencies will be distributed
and monitored through management groups set up by those agencies.
However, the entire Project (including international contributions)
will be subject to the DOE critical decision process incorporating a
CD-2 approval of its baseline cost and schedule and a CD-3 approval
for moving forward with construction.

The international DUNE Project will be responsible for monitoring and
reporting on the status of all contributions to the Project,
independent of their funding source, at least to the level of
subproject milestones tied to the critical decision process.  Non-DOE
partners will report progress to the Project through detailed
milestones related to their deliverables as defined in formal
Memoranda of Understanding (MOU).
