\subsection{Reference Design Optimization}
\label{sec:detectors-fd-ref-optimization}

%The reference design for the DUNE far detector is designed to meet the
%performance requirements. 
Considerations of physics reach as well as
cost, schedule and risk enter the process of optimizing the values for
detector design parameters such that they meet the performance requirements. 
Ideally one would like to estimate the
asymptotic far-future performance of a full simulation, reconstruction
and analysis chain, replicated for each design parameter choice and
then choose the values that maximize sensitivity while minimizing the
cost and producing timely physics results. While the GEANT4 simulation
is fairly mature, it needs to be tuned to data from the 35-t prototype
and MicroBooNE so that realistic signal and noise modeling, which are
inputs to the optimization procedures below, can be obtained. 
\fixme{So one can or cannot estimate the asymptotic far-future performance?}
The
reconstruction tools are under development (see
Section~\ref{sec:detectors-sc-physics-software}) and thus, physics
sensitivity is currently optimized using estimates of detector
performance that are input to the Fast Monte Carlo. The DUNE
collaboration plans to establish a detector performance optimization
task force to review various possible detector optimizations in light
of the new collaboration and project organizations.  This section
briefly outlines the considerations and procedures that have been and
will be used to optimize these parameters: the wire pitch, wire
angle, wire length and  maximum drift length.  The %parameters of the 
wire length, angle and pitch are directly related to the
APA dimensions, as discussed in
Section~\ref{sec:detectors-fd-ref-tpc}, and the APA dimensions are
constrained by the needs of manufacturing, storage, transport and
assembly.
\fixme{We need to decide if collaboration/project capitalized or not}

\subsubsection{Wire Pitch}

The spacing between neighboring sense wires in the APAs is an
optimizable parameter.  In principle it is freely adjustable for all
three wire planes, though to minimize the anisotropy of the detector
response, similar wire pitch should be chosen in all three planes.
The choice of a $\sim$5~mm pitch is documented in~\cite{docdb-3407}.  
The pitch of the grid plane wires is less
important as they are not instrumented, though the grid wires do
shadow the optical detectors and therefore should not be made with too fine a pitch.
\fixme{check edit above - `too fine' meant too fine a pitch, right?}

The signal-to-noise ratio is expected to be proportional to the wire
spacing, assuming that the noise on a channel is not impacted by the
presence of nearby wires, and that the signal is divided among the
available channels.  Thermal noise and uncorrelated electronics noise
satisfy these conditions.  Coherent noise is a special case ---
filters may be applied either online or offline to reduce its impact.

The signal-to-noise ratio requirement is set so that zero-suppression
can function without elaborate noise filtering.  A high signal-to-noise 
ratio %has the added benefit of improving 
improves pattern-recognition
performance, calorimetric PID performance and $dE/dx$-based
$e-\gamma$ separation.  %A large signal-to-noise ratio 
It is also important %in
for detecting sub-MIP signals, such as nuclear de-excitation photons, and
it is important when adding up the energy of hits on the edges of
showers or on the ends of stubs initiated by supernova neutrino
interactions.  

The signal-to-noise ratio is expected to be higher in
the collection plane than in the two induction planes.  The need to
deconvolve the bipolar signals while filtering noise in the induction
planes means that the collection plane will be the most reliable in
performing $dE/dx$ measurements, though for tracks that travel in a
plane containing a collection wire and the electric field, the
induction planes will be critical for recovering PID efficiency.
Reducing the spacing between wires will have an adverse impact on the
detector performance parameters that depend on the signal to noise,
with the effect seen more prominently in the induction-plane data.

In a detector with fine wire pitch, one can of course approximate a
coarser wire spacing by summing the responses on neighboring wires.
The random components of the noise will average with an RMS that grows not linearly, but 
as the square root of the number of wires. % instead of linearly.
Instead of summing signals, a more optimized approach is to deconvolve
the signal in both time and wire number, since %as 
diffusion and the field
response of the detector will cause signals to be shared among
neighboring wires.  If the diffusion characteristics are known, it is possible to either attempt
a deconvolution %may be attempted 
or perform fits with known smeared widths to expected distributions,
% can be performed, 
thereby recovering some of the
spatial resolution lost to diffusion, \fixme{where the recovery will be?} limited by noise.
\fixme{check above edit; I think the fit is smeared not the distribution; was ambiguous}

The main benefit of a finer wire pitch is the ability to obtain higher resolution
measurements of the ionization density left by events in the detector.
The separation of electrons from photons using the $dE/dx$ measured in
the initial part of an electromagnetic shower is described in
\anxreco.  The first 2.5~cm of a shower is the most important, since %as
subsequent showering stages have not yet taken place, leaving one MIP
for an electron and two for a photon conversion to two electrons,
though sometimes the subsequent showering starts earlier.  As the
first hit cannot be used to measure $dE/dx$ (since it is not known
where in the volume of argon viewed by that wire the track started \fixme{viewed by that wire? awkward}),
the second and subsequent hits must be used.  But the shower can be
aligned unfavorably along the wires of one view or another, resulting
in few usable hits.  If no hits are useful, then the $dE/dx$ method
cannot be used.  Reducing the spacing between wires in all three views
increases the precision of the measurement of the initial part of the
shower for $e-\gamma$ separation purposes.  The current study
described in \anxreco\, however, only uses the collection plane wires; % and
thus with a more optimal strategy, some of the efficiency that is lost
with 5-mm wire spacing compared with 3-mm spacing can be recovered by
examining the other two views.

Separation of multiple close tracks is improved with more closely
spaced wires.  The position resolution of hits is expected to be much
better along the drift direction than in either of the axes
perpendicular to it, as the product of the readout digitization-sample-time spacing \fixme{check
hyphening; it was hard to parse as separate words} 
%times 
and the drift velocity is much smaller than the wire spacing.  As
long as the tracks %which are desired to 
that should be separated from one another
travel at %with 
an angle with respect to the plane parallel to the wires, \fixme{travel parallel to the wires, period?}
then the fine time resolution will help with the pattern recognition
even if the wire pitch is broad.

The reconstruction of short tracks, such as low-energy protons ejected
by the struck nucleus at the primary vertex of a neutrino scattering
event, is improved with higher spatial resolution.  Finer wires also
allows for a more precise measurement of the distances between the
primary vertex and photon conversion points, which is the other
component of $\pi^0\rightarrow\gamma\gamma$ separation from electrons.
Topological identification of two EM showers and their displacement
from the primary vertex is expected to provide a factor of ten to
twenty in NC background rejection while retaining at least 90\%
efficiency for $\nu_e$CC events.  A hand-scan study comparing liquid
argon TPC detector performance in topological separation of NC events
from $\nu_e$CC events in a detector with 5-mm wire pitch and a
detector with 10-mm wire pitch~\cite{2008-hand-scan} showed no
degradation in performance with the coarser %wider 
wire pitch.  Automated
topological selection has yet to be developed, though it is
anticipated that this finding will remain true, %for automated selection as well, 
given that the radiation length in liquid argon is typically
$\sim$30 times the typical wire spacing.

%To balance against 
The gains in physics from a finer wire pitch must be balanced against the increased cost of the
electronics and online computing resources needed to read out the
additional wires. % if a finer pitch is chosen.  
In addition, the additional cold electronics
components would likely create a higher heat load in
the liquid and %is anticipated from .  M
manufacturing the APAs would likely %is expected to 
take longer %if
due to the higher density of wires. %more wires are to be installed and wrapped.

The spacing between the planes is customarily chosen to be similar to
the spacing between the wires within the planes, though this, too, is an
optimizable parameter.  Narrowing this spacing improves the sharpness
of the signals (in time) in both the induction and collection planes,
though electronics shaping and diffusion will limit the signals' ultimate sharpness. %how sharp the
%signals eventually end up being.  
These functions can be deconvolved,
though deconvolution and noise filtering produces artifacts in the
signals.  Studies varying the spacing between the planes can be
performed to estimate the impact on two-hit time resolution. % in time.

%%%%%%%%%%%  HERE!!!!!


Future studies will include the estimation of the difference in
performance of $dE/dx$-based $e-\gamma$ separation for events in which
the showers are not identified topologically as candidates for having
come from $\pi^0$ decay using automated tools.  
\fixme{I can't parse this: is it the difference in the ability to identify (blah) using hand-scans versus
automated tools? No, it's the diff betw dE/dx vs topo... Please clarify}
%
Kinematic properties
of the events, such as the energies and angles of the photons, will \fixme{allow correlation of?}
correlate the performance of $dE/dx$-based separation and topological
separation, so a full study using both techniques together will
provide the most information.  Ideally an end-to-end analysis of
neutrino scatters and the impact of particle mis-ID rates and
detection efficiencies for short tracks can be propagated to the
oscillation parameter sensitivities, but more realistically, the
effect of changing the wire pitch will be most visible in the tracking
and particle ID efficiencies.

\fixme{It's not efficient for me to spend the time to try to parse the previous pgraph. Please clarify! Anne}

\subsubsection{Wire Angle}
\label{v4:fd-ref-wireangle}

Like the wire pitch, the choice of the angles of the induction-plane
wires \fixme{relative to each other?} affects the physics performance of the detector.  Because the
wires wrap from one side of each APA to the other and back, a discrete
ambiguity is added to the continuous one \fixme{which is what?}.
% of not knowing where along a wire charge was deposited  corresponding to a measured hit on a
%channel.  The more times a wire wraps, the more choices arise for this
%ambiguity.
Given a measured hit on a channel, it is not known at what point on the wire the charge was
deposited; he more times a wire wraps, the more choices arise for this
ambiguity.

Reconstruction of 3D objects based on 2D
 data (channel number vs. time) requires associating hits in one plane
with those in at least one other. %another, or ideally, all three.  
If two wires cross only
in one place, %then there is no 
the ambiguity is removed once the hit is associated in the
two views.  If the wires cross more than once, three \fixme{views?} are required in order to
break the ambiguity of even isolated hits.

This association can most easily be done using the arrival time of the
hits.  If the time of a hit is different from that of all other hits
in the event, then the association is easy.  In more complex cases,
where dense showers produce many hits at similar drift distances,
misassociation can happen.  In this case, %If misassociation occurs, 
%the wrong choice can be made for the discrete ambiguity, 
the discrete ambiguity makes it possible to displace a reconstructed
charge deposition by multiple meters from its true location.  In the
case that the $U$ and $V$ angles are the same and the number of times
a wire wraps around an APA exceeds one, then even a single isolated
hit can be ambiguous.  A small difference in the $U$ and $V$ angles
breaks this ambiguity, though misassociation still occurs in events
with multiple nearby hits close in time.  The use of clustering methods
assists in obtaining the correct ambiguity choices for hits in dense
environments.

Reducing the wire angle reduces the number of crossings, but does not
eliminate the possibility of misassociating hits %.  
%Misassociation of hits can still happen, though, 
in events with
multiple hits simultaneously arriving.  Reducing the angle also makes
the features in the views match each other more easily as their shapes
will be more similar.
\fixme{Reducing the angle also assimilates the shapes of features in the different views
making it easier to correlate them.?}
The angle
chosen for the DUNE far detector reference design ensures that no
induction wire crosses any collection wire more than once.

On the other hand, reducing the wire angle worsens the resolution of
3D reconstruction of hits in the vertical direction. \fixme{because the
number of hits available for $dE/dx$ separation of electrons from
photons degrades for vertically-going showers?}
It also worsens the resolution on the measured separation between the
primary vertex and photon conversion points, though as pointed out
above, the radiation length is much longer than the wire spacing.  
\fixme{can you get rid of repetitive next sentence?} 
The
number of hits available for $dE/dx$ separation of electrons from
photons degrades for vertically-going showers if the wire angle is
reduced.  A parametric study of a figure-of-merit based on the
measurability of the two photon-conversion lengths as a function of
the induction-plane wire angles is provided
in~\cite{wire-orientation}.

It was estimated that the loss in resolution from even a small fraction
of grossly misplaced hits %misassociated and placed meters away from their true locations 
was worse than the resolution loss in one
dimension\cite{docdb-8981}, though studies have yet to be performed
to estimate the impact on event-detection efficiency,
particle ID performance, and energy resolution are more tied to the
final physics sensitivity of the DUNE far detector.  \fixme{can't parse previous sentence} %Work on
Advances in algorithms to break ambiguities in complex environments can also allow
for steeper wire angles.

The fact that the channel count must be an integer multiple of 128 in
an APA also constrains the wire angles as functions of the APA frame
dimensions, though a procedure to find the proper arrangement of
channels that most closely approximates the optimized wire pitches and
angles and meets the channel count constraint may cause small
deviations in the parameters.

\subsubsection{Wire Length}

The length of the collection-plane wires is determined by the APA
dimensions (or vice-versa), and the length of the induction-plane
wires is determined by their angle \fixme{relative to...?}  and the APA dimensions.  The APA
dimensions %have other constraints on them, such as 
are largely constrained by transport and handling needs %the need to transport them, over the ground, lower them down the shaft and handle them underground.  
%Long APA frames also suffer from 
as well as potential stiffness and production cost issues as they get larger.
%and may require additional cost in order to procure the members of the
%required length and stiffness.  
%Longer wires have higher capacitance
%than shorter wires and thus have more noise.  
Capacitance and thus noise increase with wire length; this effect would likely not
be masked by electronics noise since the 
cold electronics is
expected to have very low noise.  Therefore, in order to meet the
signal-to-noise requirement with a finer wire spacing, the wires may
need to be made shorter.

On the other hand, longer wires %also 
lower the cost of the detector, as fewer electronics
channels and APA frames --- and winding time and materials --- are needed to
instrument a given volume of liquid argon.

It is anticipated that much of the work needed to study the impact of
the wire pitch will inform the wire length choice due to its impact on
the signal-to-noise ratio.

\subsubsection{Maximum Drift Length}

A longer drift length is advantageous for several reasons:
\begin{itemize}
\item It reduces detector cost (this is the driving consideration).
\item It assigns more liquid argon to be read out per channel, reducing the APA count and the channel count.
\item It increases the fraction of liquid argon that is fiducial. Fiduciality cuts need to be made around the APA planes to ensure containment and to minimize the impact of dead argon inside the APA
planes; fewer APAs leads to fewer fiduciality cuts and thus to higher fiducial volume.
\end{itemize}

\fixme{The above is intended to replace next pgraph; see what you think}
The maximum drift length is another parameter that is optimizable.  In
this case, the driver for longer lengths is the cost of the detector.
A longer drift length assigns more liquid argon to be read out by any
given channel, reducing the APA count and the channel count.  A longer
drift length also increases the fraction of liquid argon that is
fiducial.  Fiduciality cuts will need to be made around the APA planes
to ensure containment and that the impact of dead argon inside the APA
planes is minimized.  Fewer APAs reduces the amount of non-fiducial
liquid argon.  The detector can be sized to have the desired fiducial
volume, but it must be larger if the maximum drift length is smaller.

A longer drift length however increases the impact of the electron
lifetime, 
\fixme{increases the NEGATIVE impact of longer e lifetime?}
as charge \fixme{electrons (only neg charge)} will attach on impurities as it drifts towards the
anode plane.  This couples the requirement on the signal-to-noise
ratio with the electron lifetime requirement and the dynamic range
requirement, given the drift length.  If the drift length were
shorter, then the electron lifetime would be less critical. \fixme{This makes it sound like
you want a shorter e lifetime. It looks like we want greater than 3 ms, not less. Please clarify}

Increasing the drift length also degrades position resolution due to
diffusion, where the spread of a drifting packet of charge increases in
proportion to the square root of the drift time.  Charge deposited
near the APAs remains well measured, though charge deposited near the
CPAs will suffer from both attenuation and diffusion, lowering the
signal-to-noise ratio.  Small-signal detection efficiencies and PID
performance may decrease for events near the CPAs.
%
Sophisticated reconstruction and analysis algorithms can be used to
recover resolution that is thus lost, but these are limited by noise.\fixme{the algorithms are limited
by noise? Or the final recovery of resolution is limited by noise?}
Simulation studies in advance of CD-2 will %study 
address the impact of
diffusion and noise on the particle reconstruction performance.
